---
title: "HsAID"
author: "Chandler Sutherland"
date: '2022-08-19'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(gridExtra)
```

Read in the excel file 
```{r}
df <- readxl::read_xlsx(path="C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E9\\Colony_count_excel\\08182022_HsAID_Rosetta2_pLysS.xlsx")

#remove too many or too few colonies
#not doing now since hand wavey test anyway and would lose all rif data 

TITC <- df %>% filter(count=='TITC') %>% count()
TNTC <- df %>% filter(count=='TNTC') %>% count()
paste(TITC$n, "plates removed for insufficient colonies to count, ", TNTC$n, "plates removed for too many colonies to count.")
df <- df %>% filter(count!='TITC' & count!='TNTC')
df$count <- as.numeric(df$count)

split <- df %>% split(f=list(df$strain))
R2 <- split[[1]]
R2_p <- split[[2]]

constructs <- list(R2 = R2, R2_p = R2_p)
```

#Since I have tech reps, I have to average them before doing my typical median thing 
```{r}
#average tech reps, then calculate mut freq for each one
R2_avg <- constructs[[1]] %>% group_by(plate, induction, culture) %>% summarize(tech.mean=mean(CFU))
lb_kan <- R2_avg[which(R2_avg$plate == 'LB_KAN'),]$tech.mean
lb_rif_kan <- R2_avg[which(R2_avg$plate == 'LB_RIF_KAN'),]$tech.mean
mut_freq = lb_rif_kan/lb_kan
R2_mut <- R2_avg %>% group_by(induction, culture) %>% summarize(n=n())
R2_mut$mut_freq <- mut_freq
R2_mut

R2_p_avg <- constructs[[2]] %>% group_by(plate, induction, culture) %>% summarize(tech.mean=mean(CFU))
lb_cam_kan <- R2_p_avg[which(R2_p_avg$plate == 'LB_CAM_KAN'),]$tech.mean
lb_rif_cam_kan <- R2_p_avg[which(R2_p_avg$plate == 'LB_RIF_KAN_CAM'),]$tech.mean
mut_freq = lb_rif_cam_kan/lb_cam_kan
R2_p_mut <- R2_p_avg %>% group_by(induction, culture) %>% summarize(n=n())
R2_p_mut$mut_freq <- mut_freq
R2_p_mut
```

```{r}
R2 <- R2_mut %>% group_by(induction) %>% summarize(median.mut_freq = median(mut_freq), 
                                                   mean.mut_freq = mean(mut_freq),
                                                   sd.mut_freq = sd(mut_freq), 
                                                   n.mut_freq = n())
R2$strain <- c('Rosetta2', 'Rosetta2')
R2_p <- R2_p_mut %>% group_by(induction) %>% summarize(median.mut_freq = median(mut_freq), 
                                                       mean.mut_freq = mean(mut_freq),
                                                       sd.mut_freq = sd(mut_freq), 
                                                       n.mut_freq = n())
R2_p$strain <- c('Rosetta2_pLysS', 'Rosetta2_pLysS')
median.mut_freq <- rbind(R2, R2_p)
median.mut_freq
```
The induced mut freq in R2_pLysS is greater than uninduced, but has pretty big sd 

Culture 3 in pLysS is a bit of an outlier, and some dried up plates on the uninduced side so many be distorting 
```{r}
R2_p_filt <- R2_p_mut %>% filter(culture != 3) %>% group_by(induction) %>% summarize(median.mut_freq = median(mut_freq), 
                                                       mean.mut_freq = mean(mut_freq),
                                                       sd.mut_freq = sd(mut_freq), 
                                                       n.mut_freq = n())

R2_p_filt

```
Ok, that brings down the sd to a normal amount


With this information, is there a statistical difference between induced and uninduced in each condition? 
```{r}
#t test time 

R2_t <- t.test(mut_freq ~ induction, data=R2_mut)$p.value
R2_p_t <- t.test(mut_freq ~ induction, data=R2_p_mut)$p.value
R2_p_filt_t <- t.test(mut_freq ~ induction, data=R2_p_mut %>% filter(culture != 3))$p.value
print(paste("HsAID does not show significant increase in mutation frequency in either condition, with a p value of ", R2_t, "in Rosetta2 and ", R2_p_t, "in Rosetta2 + pLysS. Removing culture 3 does not change significance, with a p value of", R2_p_filt_t))
```

```{r}
#This forces uninduced before induced
R2_mut$induction <- factor(R2_mut$induction, levels = c('uninduced', 'induced'), ordered=TRUE)
R2_p_mut$induction <- factor(R2_p_mut$induction, levels = c('uninduced', 'induced'), ordered=TRUE)

#plot as boxplots 
R2_plot <- ggplot(R2_mut, 
       aes(x=reorder(induction,-mut_freq), y=mut_freq, fill=induction)) + 
  geom_boxplot() + 
  geom_jitter() +
  labs(x = 'induction', y='mutation frequency', title="HsAID+Rosetta2")

R2_p_plot <- ggplot(R2_p_mut, 
       aes(x=reorder(induction, mut_freq), y=mut_freq, fill=induction)) + 
  geom_boxplot()+
  geom_jitter()+
  labs(x = 'induction', y='mutation frequency', title="HsAID+Rosetta2 pLysS")

#without culture 3
R2_p_filt_plot <- ggplot(R2_p_mut %>% filter(culture != 3), 
       aes(x=reorder(induction, mut_freq), y=mut_freq, fill=induction)) + 
  geom_boxplot()+
  geom_jitter()+
  labs(x = 'induction', y='mutation frequency', title="HsAID+Rosetta2 pLysS")

R2_p_filt_mut <- R2_p_mut %>% filter(culture != 3)
R2_mut$strain <- 'Rosetta2'
R2_p_filt_mut$strain <- 'Rosetta2_pLysS'

R2_mut$induction <- factor(R2_mut$induction, levels = c('uninduced', 'induced'), ordered=TRUE)
R2_p_filt_mut$induction <- factor(R2_p_filt_mut$induction, levels = c('uninduced', 'induced'), ordered=TRUE)

ggplot(rbind(R2_mut,R2_p_filt_mut), aes(x=induction, y=mut_freq, fill=induction, label=culture)) + 
  geom_boxplot() + 
  facet_wrap(~strain) +
  geom_jitter() +
  labs(x = 'induction', y='mutation frequency', title="HsAID")
```

What if I remove rep3 that seems to be introducing a lot of noise to R2_pLysS?

```{r}
ggplot(rbind(R2_mut, R2_p_mut %>% filter(culture != 3)), aes(x=induction, y=mut_freq, fill=induction, label=culture)) + 
  geom_boxplot() + 
  facet_wrap(~strain) +
  geom_point() +
  geom_text()+
  labs(x = 'induction', y='mutation frequency', title="HsAID")
```
Looks tighter, but still not statistically significant.  


Because I did try this before, I can load in some previous HsAID and see a true bio rep 
```{r}
df2 <- readxl::read_xlsx(path="C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E9\\Colony_count_excel\\02042022_HsAID_Rosetta2p.xlsx")
df2 <- df2 %>% filter(plate != 'LB')

#all tech reps 
df2_avg <- df2 %>% group_by(plate, induction) %>% summarize(tech.mean=mean(CFU))
lb_cam_kan <- df2_avg[which(df2_avg$plate == 'LB_CAM_KAN'),]$tech.mean
lb_rif_cam_kan <- df2_avg[which(df2_avg$plate == 'LB_RIF_CAM_KAN'),]$tech.mean
mut_freq = lb_rif_cam_kan/lb_cam_kan
df2_mut <- df2_avg %>% group_by(induction) %>% summarize(n=n())
df2_mut$mut_freq <- mut_freq
df2_mut$culture <- 7
df2_mut$strain <- 'Rosetta2_pLysS'


ggplot(rbind(df2_mut, R2_p_filt_mut), aes(x=induction, y=mut_freq, fill=induction, label=culture)) + 
  geom_boxplot() + 
  facet_wrap(~strain) +
  geom_point() +
  geom_text()+
  labs(x = 'induction', y='mutation frequency', title="HsAID")
```

Wow, looks really clean. Nice. 
