---
title: "controls"
author: "Chandler Sutherland"
date: '2022-07-29'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
```

## Functions
Read in the excel file 
```{r}
df <- readxl::read_xlsx(path="C:\\Users\\chand\\Box Sync\\Krasileva_Lab\\Research\\chandler\\Krasileva Lab\\E9\\Colony_count_excel\\08062022_controls_Rosetta2.xlsx")

#remove too many or too few colonies
#not doing now since hand wavey test anyway and would lose all rif data 
##add if statement to do it for me
#TITC <- df %>% filter(CFU=='TITC') %>% count()
#TNTC <- df %>% filter(CFU=='TNTC') %>% count()
#paste(TITC$n, "plates removed for insufficient colonies to count, ", TNTC$n, "plates removed for too many colonies to count.")
#df <- df %>% filter(CFU!='TITC' & CFU!='TNTC')
#df$CFU <- as.numeric(df$CFU)

split <- df %>% split(f=list(df$construct, df$strain))
HsAID <- split[[1]]
HsApo <- split[[2]]
HsCDD <- split[[3]]
PmCDA <- split[[4]]
HsApo_p <- split[[6]]

constructs <- list(HsAID = HsAID, HsApo = HsApo, HsCDD = HsCDD, PmCDA = PmCDA, HsApo_p = HsApo_p)
```

```{r}
summary_ize <- function(x){
  #this function takes in a dataframe, and outputs the summary I am interested in. 
  summary <- x %>% group_by(plate,induction)%>%
    summarize(CFU.mean=mean(CFU),
              CFU.median=median(CFU),
              CFU.sd=sd(CFU),
              CFU.n=n(),
              CFU.se=CFU.sd/sqrt(CFU.n))
  summary
}


construct_summary <- lapply(constructs, summary_ize)
construct_summary %>% names()
```

```{r}
#The mutation frequency for a given condition is the median CFU RIF R/LB_KAN 
HsAID_summary <- construct_summary$HsAID 
construct_summary$HsApo_p

#This function calculates the mutation frequency. Caution: not vague enough to handle different plate names, so have to do pLysS by hand 
mut_freq <- function(x){
  un_mut_freq <- x[x$plate=='LB_RIF_KAN'& x$induction == 'uninduced',"CFU.median"][[1]]/
    x[x$plate=='LB_KAN'& x$induction == 'uninduced', "CFU.median"][[1]]
  
  in_mut_freq <- x[x$plate=='LB_RIF_KAN'& HsAID_summary$induction == 'induced', "CFU.median"][[1]]/
    x[x$plate=='LB_KAN'& x$induction == 'induced', "CFU.median"][[1]]
  list(un_mut_freq=un_mut_freq, in_mut_freq=in_mut_freq)
}

mut_freqs <- lapply(construct_summary, mut_freq)
do.call(rbind, mut_freqs) 

#for apobec_p 
HsApo_summary <- construct_summary$HsApo_p
un_mut_freq <- HsApo_summary[HsApo_summary$plate=='LB_RIF_CAM_KAN'& HsApo_summary$induction == 'uninduced',"CFU.median"][[1]]/
    HsApo_summary[HsApo_summary$plate=='LB_CAM_KAN'& HsApo_summary$induction == 'uninduced', "CFU.median"][[1]]

in_mut_freq <- HsApo_summary[HsApo_summary$plate=='LB_RIF_CAM_KAN'& HsApo_summary$induction == 'induced',"CFU.median"][[1]]/
    HsApo_summary[HsApo_summary$plate=='LB_CAM_KAN'& HsApo_summary$induction == 'induced', "CFU.median"][[1]]
```
The only case where the induced mutation frequency is > the uninduced is HsCDD. HsAID the uninduced has a higer baseline mutation frequency though. Caveats: combining two different dilutions of RIF colonies, only two reps of those, almost all are TITC. 

With this information, is there a statistical difference between induced and uninduced in each condition? 
```{r}
##has to be a cleaner way to do this 
##also is this the correct kind of t test to do?
apo_LB_KAN <- HsAPOBEC %>% filter(plate == 'LB_KAN')
apo_LB_KAN_t <- t.test(CFU ~ induction, data=apo_LB_KAN)$p.value
apo_LB_RIF_KAN <- HsAPOBEC %>% filter(plate == 'LB_RIF_KAN')
apo_LB_RIF_KAN_t <- t.test(CFU ~ induction, data=apo_LB_RIF_KAN)$p.value

print(paste("The HsAPOBEC uninduced vs induced CFU/mL on LB KAN mean is not significantly different (barely) with a p value of", apo_LB_KAN_t, ". It is also an insignificant difference on LB RIF KAN, with a p value of", apo_LB_RIF_KAN_t))
```

```{r}
cda_LB_KAN <- PmCDA1 %>% filter(plate == 'LB_KAN')
cda_LB_KAN_t <- t.test(CFU ~ induction, data=cda_LB_KAN)$p.value
cda_LB_RIF_KAN <- PmCDA1 %>% filter(plate == 'LB_RIF_KAN')
cda_LB_RIF_KAN_t <- t.test(CFU ~ induction, data=cda_LB_RIF_KAN)$p.value
cda_LB_KAN_t
paste("The PmCDA1 CFUs/mL on LB KAN and LB RIF KAN means are significantly different, with p values of ", cda_LB_KAN_t, "and", cda_LB_RIF_KAN_t, "respectively.")
```

```{r}
ggplot(apo_LB_KAN, 
       aes(x=reorder(induction,-CFU), y=CFU, fill=induction)) + 
  geom_boxplot() + 
  geom_jitter() +
  labs(x = 'induction', y='CFUs/mL', title="HsAPOBEC on LB KAN")

ggplot(apo_LB_RIF_KAN, 
       aes(x=reorder(induction, CFU), y=CFU, fill=induction)) + 
  geom_boxplot()+
  geom_jitter()+
  labs(x = 'induction', y='CFUs/mL', title="HsAPOBEC on LB RIF KAN")

#This forces uninduced before induced
HsAPOBEC$induction <- factor(HsAPOBEC$induction, levels = c('uninduced', 'induced'), ordered=TRUE)
PmCDA1$induction <- factor(PmCDA1$induction, levels = c('uninduced', 'induced'), ordered=TRUE)

#plot
ggplot(HsAPOBEC, aes(x=induction, y=CFU, fill=induction)) + 
  geom_boxplot() + 
  facet_wrap(~plate, scales='free') +
  geom_jitter() +
  labs(x = 'induction', y='CFUs/mL', title="HsAPOBEC")

```
```{r}
ggplot(PmCDA1, aes(x=induction, y=CFU, fill=induction)) + 
  geom_boxplot() + 
  facet_wrap(~plate, scales='free') +
  geom_jitter() +
  labs(x = 'induction', y='CFUs/mL', title="PmCDA1")
```
```{r}
HsAPOBEC_summary$induction <- factor(HsAPOBEC_summary$induction, levels = c('uninduced', 'induced'), ordered=TRUE)
ggplot(HsAPOBEC_summary, aes(x=induction, y=CFU.mean, fill=induction)) + 
  geom_col() + 
  geom_errorbar(aes(ymin=CFU.mean-CFU.se, ymax=CFU.mean+CFU.se), width=.2)+
  facet_wrap(~plate, scales='free')+
  labs(x = 'induction', y='CFUs/mL', title="HsAPOBEC on Rosetta2")
```
```{r}
PmCDA1_summary$induction <- factor(PmCDA1_summary$induction, levels = c('uninduced', 'induced'), ordered=TRUE)
ggplot(PmCDA1_summary, aes(x=induction, y=CFU.mean, fill=induction)) + 
  geom_col() + 
  geom_errorbar(aes(ymin=CFU.mean-CFU.se, ymax=CFU.mean+CFU.se), width=.2)+
  facet_wrap(~plate, scales='free')+
  labs(x = 'induction', y='CFUs/mL', title="PmCDA1 on Rosetta2")
```

The standard deviation is a measure of variability. The SD of a sample is an estimate of variability of the poulation from which the sample was drawn. 

The standard error describes how much the sample mean varies from the standard deviatio of this sampling distribution, which is the SE of the estiamte of the mean. 

The LB_KAN replicates are from the same dilution series, so that is a really tight technical replicate between plating efficiencies. The LB_RIF_KAN were at least different aliquots, so that noise is from how shaken the cultures were, my pipetting accuracy, and plating efficiency. And some time in between sampling (30 minutes at RT). So I am interested in the SD to look at differences across these conditions, but more generally the SE to see how close I am to my population mean. Biological replicates would be different cultures. 

```{r}
count_stats <- df %>% group_by(construct, plate, induction, dilution_factor)%>%
  summarize(count.mean=mean(count), 
            count.sd=sd(count)) %>%
  arrange(plate) %>% filter(count.mean!=0)

#to compare the SDs between plates or constructs, the F test will tell me if there is statistically significant difference between my two variances. 
LB_KAN <- count_stats[count_stats$plate == 'LB_KAN',][['count.sd']]
LB_RIF_KAN <- count_stats[count_stats$plate == 'LB_RIF_KAN',][['count.sd']]
var.test(LB_KAN, LB_RIF_KAN)

#looks to be not statistically significant, but this test isn't very powerful. Try ANOVA also. 
```


